---
title: "Strip"
description: |
  What makes a good graph? One feature is certainly the efficient communication of information. In this lecture we try to apply Tufte's idea of 'data-ink-ratio'.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      layout="l-body-outset",
                      dpi=300)
if(!require("pacman"))install.packages("pacman")
pacman::p_load("tidyverse", "ggthemes")
```

*Note: This first part is in large parts just a recreation of the examples given by Martin Rolfs in Matlab.*

# Data

Just renaming variables, to fit my personal preferences.

```{r}
data(iris)
iris <- rename_all(iris, tolower) %>%
  rename_all(~str_replace(., fixed("."), "_"))
```

# Scatter Plot

This is our basic plot we want to improve.

```{r}
scatter_orginal <- ggplot(iris, aes(sepal_length, petal_length)) + geom_point()
plot(scatter_orginal)
```

Clearly we should code the species.

```{r}
scatter <- ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  NULL
plot(scatter)
```

I just cannot stand the standard grey backround.

```{r}
scatter <- scatter +
  theme_minimal() +
  NULL
plot(scatter)
```

Both axes are on the same scale, we could reflect that.

```{r}
scatter <- scatter + 
  coord_fixed() +
  NULL
plot(scatter)
```

The breaks are not symetrical.

```{r}
round_breaks <- function(lim)ceiling(seq(lim[[1]], lim[[2]], by = 1))
scatter <- scatter +
  scale_x_continuous(breaks = round_breaks) +
  scale_y_continuous(breaks = round_breaks) +
  NULL
plot(scatter)
```

The legend does need a lot of space.

```{r}
iris_cluster <- iris %>%
  group_by(species) %>%
  summarise(sepal_length = mean(sepal_length),
            petal_length = mean(petal_length))

scatter <- scatter + 
  geom_text(data = iris_cluster,
            mapping = aes(sepal_length,
                          petal_length,
                          label = species,
                          color = species),
            nudge_x = 1.5,
            inherit.aes = FALSE) +
  theme(legend.position = "none") +
  NULL
plot(scatter)
```

Adding a marginal distribution comes essentially for free.

```{r}
scatter <- scatter +
  geom_jitter(aes(x = 4, y = petal_length), width = 0, shape = 96, size = 10, alpha = .5) +
  geom_jitter(aes(x = sepal_length, y = 0), height = 0, shape = 96, size = 10, alpha = .5) +
  NULL
plot(scatter)
```

We gain a lot of space by flipping the plot.

```{r}
scatter <- scatter +
  coord_flip() +
  NULL
plot(scatter)
```

Ok, proper labels are neccesary.

```{r}
scatter <- scatter +
  xlab("Sepal lenght (cm)") + 
  ylab("Petal lenght (cm)") +
  NULL
plot(scatter)
```

Personally I all ready like this graph but lets get more tufteesque. Tufte would probably replace the backround grid by axis ticks.

```{r}
scatter <- scatter + 
  theme_void() +
  theme(axis.ticks = element_line(),
        axis.ticks.length = unit(0.05, "cm"), 
        axis.text = element_text(size = rel(0.8)),
        legend.position = "none",
        axis.text.x = element_text(margin = margin(t = 5), vjust = 0),
        axis.text.y = element_text(margin = margin(r = 5), hjust = 0),
        axis.title = element_text(),
        axis.title.y = element_text(angle = 90, margin = margin(r = 5)),
        axis.title.x = element_text(angle = 0, margin = margin(t = 5)),) + 
  NULL
plot(scatter)
```

I am unsure which colors tufte might use but probably something more muted.

```{r}
avg_color <- function(hexs){
  rgbs <- map(hexs, col2rgb)
  rescaled <- do.call(cbind, rgbs)/255
  avg <- rowMeans(rescaled)
  do.call(rgb, as.list(avg))
}
avg_traffic <- avg_color(tableau_color_pal("Traffic")(3))
scatter <- scatter +
  theme(axis.ticks = element_line(colour = avg_traffic),
        text = element_text(colour = avg_traffic)) + 
  scale_colour_tableau("Traffic") +
  NULL
plot(scatter)
```

Tufte dislikes displays, so he often presents digital representation with a "non-white" backround that resembles paper.

```{r}
scatter <- scatter +
  theme(plot.background = element_rect(fill = "#FFFAF5"),
        plot.margin = unit(c(1,1,1,1), "cm"))
plot(scatter)
```

## Result

```{r, echo=FALSE, layout="l-screen-inset"}
plot(scatter_orginal)
plot(scatter)
```

# Boxplot

```{r}
iris_wide <- iris %>%
  pivot_longer(c(sepal_length,
                                petal_length,
                                sepal_width,
                                petal_width),
                              names_to = "measure",
                              values_to = "cm")
box <- ggplot(iris_wide, aes(measure, cm, color = species)) +
  geom_boxplot() +
  NULL
plot(box)
```

There are multiple critics --- one of them tufte --- of the boxplot, the main argument is that it hides the raw data. So I revert to simple points, since it is a small dataset.

```{r}
point <- ggplot(iris_wide, aes(measure, cm, color = species, group = species)) +
  geom_point() +
  NULL
plot(point)
```

Okay the overlap between the groups hides much information.

```{r}
point <- ggplot(iris_wide, aes(measure, cm, color = species, group = species)) +
  geom_point(position = position_dodge(width = 1)) +
  NULL
plot(point)
```

Still a lot of overplotting, so I arrange them ordered by their average size, measured as the mean of z-values on all four variables per species.

```{r}
iris_wide <- iris %>%
    group_by(species) %>%
  mutate(mean_size = rowMeans(cbind(scale(petal_length),
                                    scale(petal_width),
                                    scale(sepal_length),
                                    scale(sepal_width)))) %>% 
  arrange(mean_size) %>% 
  rownames_to_column("id") %>% 
  pivot_longer(c(sepal_length,
                                petal_length,
                                sepal_width,
                                petal_width),
                              names_to = "measure",
                              values_to = "cm") %>% 
  ungroup() %>% 
  mutate(id = factor(as.numeric(id)))
point <- ggplot(iris_wide, aes(measure, cm, color = species, group = species)) +
  geom_point(position = position_jitterdodge(jitter.height = 0, seed = 1234)) +
  labs(caption = "Points are ordered according to average ranked size.") +
  NULL
plot(point)
```

Again, replace the backround.

```{r}
point <- point +
  theme_minimal() +
  NULL
plot(point)
```

I think this plot mixes two dimensions on the x-axes, length vs. width and sepal vs. petal.

```{r}
iris_wide2 <- iris_wide %>%
  separate(measure, c("leaf", "measure"), sep = "_")
point <- ggplot(iris_wide2, aes(species, cm, color = species, group = id)) +
  geom_point(position = position_dodge(width = .8), alpha = .7, size = .7) +
  facet_grid(measure~leaf) +
  theme_minimal() +
  NULL
plot(point)
```

While the ordering provides interesting information, it makes it a bit harder to judge the distribution, so I add a minimal boxplot.

```{r}
boxplot_stats <- function(x, coef = 1.5){
  raw <- boxplot.stats(x, coef)
  tibble(whisker_position = factor(c("upper", "upper", "lower", "lower")),
         whisker = c(raw$stats[-3]),
         median = raw$stats[3])
}
iris_summary <- iris_wide2 %>%
  group_by(species, leaf, measure) %>%
  summarise(boxplot = list(boxplot_stats(cm))) %>% 
  unnest(boxplot) %>% 
  unite("whisker_position", species, whisker_position, remove = FALSE)
point <- point +
  geom_point(aes(species, median, color = species),
                   data = iris_summary,
                   position = position_nudge(x = .5),
             size = .7,
                   inherit.aes = FALSE) +
  geom_line(aes(species, whisker, group = whisker_position, color = species),
                   data = iris_summary,
                   position = position_nudge(x = .5),
                   inherit.aes = FALSE) +
  labs(caption = "Points are ordered along x-axes according to average ranked size.") +
  theme(plot.caption = element_text(margin = unit(c(.5, 0, 0, 0), "cm"))) +
  NULL
plot(point)
```

I think the boxplots seperate the species clearly enough, so that the legend may be remved.

```{r}
point <- point + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
plot(point)
```

Maybe we can remove the grid:

```{r}
point <- point +
theme_void() +
  theme(axis.ticks = element_line(),
        axis.ticks.length = unit(0.05, "cm"), 
        axis.text = element_text(size = rel(0.8)),
        legend.position = "none",
        axis.text.x = element_text(margin = margin(t = 5), vjust = 0),
        axis.text.y = element_text(margin = margin(r = 5), hjust = 0)) + 
  NULL
plot(point)
```

I liked the color scheme for the scatter plot.

```{r}
point <- point + geom_vline(aes(xintercept = x),
                            tibble(leaf = "petal", x = Inf),
                            col = avg_traffic,
                            alpha = .2) +
  geom_hline(aes(yintercept = x),
             tibble(measure = c("length"), x = 0L),
             col = avg_traffic,
             alpha = .2) +
  scale_colour_tableau("Traffic") +
  theme(axis.ticks = element_line(colour = avg_traffic),
        text = element_text(colour = avg_traffic),
        plot.background = element_rect(fill = "#FFFAF5"),
        panel.spacing = unit(0, "cm"),
        plot.margin = unit(c(1, 1, 1, 1), "cm")) +
  NULL
plot(point)
```

## Result

```{r, echo=FALSE, layout="l-screen-inset"}
plot(box)
plot(point)
```
