---
title: "Strip"
description: |
  What makes a good graph? One feature is certainly the efficient communication of information.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
if(!require("pacman"))install.packages("pacman")
pacman::p_load("tidyverse", "ggthemr")
```

# Data

Just renaming, to fit my personal preferences.

```{r}
data(iris)
iris <- rename_all(iris, tolower) %>%
  rename_all(~str_replace(., fixed("."), "_"))
```

# Plots

This is our basic plot we want to improve.

```{r}
ggplot(iris, aes(sepal_length, petal_length)) + geom_point()
```

Clearly we should code the species.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  NULL
```

I just cannot stand the standard grey backround.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  NULL
```

Both axes are on the same scale, we could reflect that.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  NULL
```

The breaks are not symetrical.

```{r}
round_breaks <- function(lim)ceiling(seq(lim[[1]], lim[[2]], by = 1))
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(breaks = round_breaks) +
  scale_y_continuous(breaks = round_breaks) +
  NULL
```

The legend does need a lot of space.

```{r}
iris_cluster <- iris %>%
  group_by(species) %>%
  summarise(sepal_length = mean(sepal_length),
            petal_length = mean(petal_length))

ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(breaks = round_breaks, limits = c(4, 9)) +
  scale_y_continuous(breaks = round_breaks) +
  geom_text(data = iris_cluster,
            mapping = aes(sepal_length,
                          petal_length,
                          label = species,
                          color = species),
            nudge_x = 1.5,
            inherit.aes = FALSE) +
  theme(legend.position = "none") +
  NULL
```

Adding a marginal distribution comes essentially for free.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(breaks = round_breaks, limits = c(4, 9)) +
  scale_y_continuous(breaks = round_breaks) +
  geom_text(data = iris_cluster,
            mapping = aes(sepal_length,
                          petal_length,
                          label = species,
                          color = species),
            nudge_x = 1.5,
            inherit.aes = FALSE) +
  theme(legend.position = "none") +
  xlab("Sepal lenght (cm)") + 
  ylab("Petal lenght (cm)") + 
  geom_jitter(aes(x = 4, y = petal_length), width = 0, shape = 96, size = 4, alpha = .5) +
  geom_jitter(aes(x = sepal_length, y = 0), height = 0, shape = 96, size = 4, alpha = .5) +
  NULL
```

We gain a lot of space by flipping the plot.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(breaks = round_breaks, limits = c(4, 9)) +
  scale_y_continuous(breaks = round_breaks) +
  geom_text(data = iris_cluster,
            mapping = aes(sepal_length,
                          petal_length,
                          label = species,
                          color = species),
            nudge_x = 1.5,
            inherit.aes = FALSE) +
  theme(legend.position = "none") +
  xlab("Sepal lenght (cm)") + 
  ylab("Petal lenght (cm)") + 
  geom_jitter(aes(x = 4, y = petal_length), width = 0, shape = 96, size = 4, alpha = .5) +
  geom_jitter(aes(x = sepal_length, y = 0), height = 0, shape = 96, size = 4, alpha = .5) +
  coord_flip() +
  NULL
```

Personally I all ready like this graph but one may replace the backround grid by axis ticks.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(breaks = round_breaks, limits = c(4, 9)) +
  scale_y_continuous(breaks = round_breaks) +
  geom_text(data = iris_cluster,
            mapping = aes(sepal_length,
                          petal_length,
                          label = species,
                          color = species),
            nudge_x = 1.5,
            inherit.aes = FALSE) +
  theme(legend.position = "none") +
  xlab("Sepal lenght (cm)") + 
  ylab("Petal lenght (cm)") + 
  geom_jitter(aes(x = 4, y = petal_length), width = 0, shape = 96, size = 4, alpha = .5) +
  geom_jitter(aes(x = sepal_length, y = 0),
              height = 0, shape = 96, size = 4, alpha = .5) +
  coord_flip() +
  theme_void() +
  theme(axis.ticks = element_line(),
        axis.ticks.length = unit(0.05, "cm"), 
        axis.text = element_text(size = rel(0.8)),
        legend.position = "none",
        axis.text.x = element_text(margin = margin(t = 5), vjust = 0),
        axis.text.y = element_text(margin = margin(r = 5), hjust = 0),
        axis.title = element_text(),
        axis.title.y = element_text(angle = 90, margin = margin(r = 5)),
        axis.title.x = element_text(angle = 0, margin = margin(t = 5)),) + 
  NULL
```

Now we can mute the colours a bit.

```{r}
ggplot(iris, aes(sepal_length, petal_length, color = species)) +
  geom_point() +
  theme_minimal() +
  coord_fixed() +
  scale_x_continuous(breaks = round_breaks, limits = c(4, 9)) +
  scale_y_continuous(breaks = round_breaks) +
  geom_text(data = iris_cluster,
            mapping = aes(sepal_length,
                          petal_length,
                          label = species,
                          color = species),
            nudge_x = 1.5,
            inherit.aes = FALSE) +
  theme(legend.position = "none") +
  xlab("Sepal lenght (cm)") + 
  ylab("Petal lenght (cm)") + 
  geom_jitter(aes(x = 4, y = petal_length), width = 0, shape = 96, size = 4, alpha = .5) +
  geom_jitter(aes(x = sepal_length, y = 0),
              height = 0, shape = 96, size = 4, alpha = .5) +
  coord_flip() +
  theme_void() +
  theme(axis.ticks = element_line(colour = "#E0C89D"),
        axis.ticks.length = unit(0.05, "cm"), 
        axis.text = element_text(size = rel(0.8)),
        legend.position = "none",
        axis.text.x = element_text(margin = margin(t = 5), vjust = 0),
        axis.text.y = element_text(margin = margin(r = 5), hjust = 0),
        axis.title = element_text(),
        axis.title.y = element_text(angle = 90, margin = margin(r = 5)),
        axis.title.x = element_text(angle = 0, margin = margin(t = 5)),
        text = element_text(colour = "#E0C89D")) + 
  scale_color_manual(values = c("#F5BB5C", "#D96725", "#731702")) +
  NULL
```

